---
title: "Course project"
subtitle: "Total CO2 emissions in Latin America countries"
author: "Oscar Nieto G"
date: "6/15/2022"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

Some visualizations of the behavior of the CO$_{2}$ emissions by Latin America countries are presented. The dataset was obtained from a collection of metrics maintained by [Our World in Data](https://ourworldindata.org/co2-and-other-greenhouse-gas-emissions), available at [co2-data](https://raw.githubusercontent.com/owid/co2-data/master/owid-co2-data.csv).

```{r out.width="300px", fig.align="center"}

url <- "https://as1.ftcdn.net/v2/jpg/01/23/03/48/1000_F_123034829_JYOf4x538pMkbB1HhkK90aPGOetAHZvR.jpg"
knitr::include_graphics(url)

```

## Time series plot

To represent the behavior of the CO$_{2}$ emissions on Latin America in the last 30 years or less a time series plot were generated, representing the trend by each selected country.

```{r echo = FALSE, message=FALSE}

  
library(tidyverse)
library(plotly)

df <- read.csv('https://raw.githubusercontent.com/owid/co2-data/master/owid-co2-data.csv')
lac_countries <- c("Belize", "Costa Rica", "El Salvador", "Guatemala", "Honduras", "Mexico", "Nicaragua", "Panama", "Argentina", "Bolivia", "Brazil", "Chile", "Colombia", "Ecuador",
                   "Guyana", "Paraguay", "Peru", "Suriname", "Uruguay", "Venezuela", "Cuba", "Dominican Republic",
                   "Haiti", "Puerto Rico")

df %>% filter(year > 1990, country %in% lac_countries) %>%
  select(c("iso_code", "country", "year", "co2", "co2_per_capita")) -> df_lac30

df_lac30 %>% group_by(country) %>% summarize(tot_co2 = sum(co2)) %>%
  arrange(desc(tot_co2)) %>% dplyr::pull(country) -> country

top <-  country[1:5]

df_lac30 %>% filter(country %in% top) %>%
  ggplot( aes(x=year, y=co2, color=country)) +
  geom_line(size = 0.7) + 
  scale_color_brewer(palette = "Set2") +
  labs(x="", y="Millon Tonnes", title= bquote('Annual'~CO[2]~'Emissions' ) ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5, size=20, face="bold")) -> p

plotly::ggplotly(p) 

```

## Map with the total distribution of CO$_{2}$

The following map shows the total CO$_{2}$ emitted by Latin American countries in the last 30 years.

```{r echo = FALSE, message=FALSE}

library(leaflet)
library(maps)

df_lac30 %>% group_by(country) %>% summarize(tot_co2 = sum(co2)) %>%
  arrange(desc(tot_co2)) -> tot_co2

mybins <- c(0,100,500,1000,5000,10000,50000)
mypalette <- colorBin( palette="YlOrBr", domain=tot_co2$tot_co2, na.color="transparent", bins=mybins)
bounds <- map("world", tot_co2$country, fill = TRUE, plot = FALSE)
        
        leaflet() %>%
            addProviderTiles(providers$Stamen.TonerLite,
                             options = providerTileOptions(noWrap = TRUE) ) %>%
            
            addPolygons(data = bounds, group = "country", 
                        fillColor =  ~mypalette(tot_co2$tot_co2),
                        stroke=TRUE, 
                        color = "white",
                        weight=0.3,
                        popup = paste("Country: ", bounds$names),
                        fillOpacity = 0.9,
            ) %>%
            addLegend(pal=mypalette, values=tot_co2, opacity=0.9, title = "Total Emissions (CO2)", position = "bottomleft" )

```

## Conclusions

Finally, we can highlight some key insights:

- Mexico and Brazil are the two countries that have emitted more CO$_{2}$ in Latin America in the last 30 years.
- There was an upward trend on the emissions between 1990 to 2014, after this year there was an slight decrease in the emissions of the five top countries. 
